Model Estimation

### 2019-02-11 model_estimation.Rmd
Lock down coefficients on size and distance.

tracts@data <-  tbl_df(tracts@data) %>%
  dplyr::transmute(
    geoid = GEOID,
    county_fips = substr(GEOID, 3, 5),
    borough = case_when(
      county_fips == "081" ~ "Queens",
      county_fips == "047" ~ "Brooklyn",
      county_fips == "061" ~ "Manhattan",
      county_fips == "005" ~ "Bronx",
      county_fips == "085" ~ "Staten Island",
      TRUE ~ as.character(NA)
    ),
    obesity = OBESITY,
    physact = 100 - Phys_Act,
    log_obesity = log(OBESITY),
    log_physact = log(100 - Phys_Act),
    density = Pop_Density,
    fulltime = FulltimeWork, college = CollegeDeg,
    single = Single_Percent,
    black = PctBlack, asian = PctAsian, hispanic = PctHispanic,
    other = PctNative + PctPacific,
    Pct0to17, Pct18to29, Pct65plus,
    income = NYC_Income
  )
```

@@

```{r W, echo=FALSE}
tracts.dnn <- dnearneigh(gCentroid(tracts, byid = TRUE), 0, 1.8 * 5280)
dists <- nbdists(tracts.dnn, gCentroid(tracts, byid = TRUE))
dists.inv <- lapply(dists, function(x) 1 / x)
W <- nb2listw(neighbours = tracts.dnn, glist = dists.inv, zero.policy = TRUE, style = "W")
print(W, zero.policy = TRUE)
trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
```

@@

```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
base_formula <- formula(
  ~ log(density) + log(income) +
    fulltime + college + single +
    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity
    black + asian + hispanic + other)
```

@@

```{r spatial selection}
physact_base_lm <- lm(update(base_formula, physact ~ .), data = tracts)
physact_base_sar <- lagsarlm(update(base_formula, physact ~ .),
                           data = tracts@data, listw = W, zero.policy = TRUE)
physact_base_sem <- errorsarlm(update(base_formula, physact ~ .),
                             data = tracts@data, listw = W, zero.policy = TRUE)
physact_base_sdm <- lagsarlm(update(base_formula, physact ~ .),
                           data = tracts@data, listw = W, zero.policy = TRUE,
                           type = "mixed")
test_sdmsem <- lmtest::lrtest(physact_base_sem, physact_base_sdm)
```

@@

```{r }
physact_models <- list("OLS" = physact_base_lm, "SAR" = physact_base_sar, 
                       "SEM" = physact_base_sem, "SDM" = physact_base_sdm)
screenreg(physact_models, digits = 5)

@@

```{r residuals}
tracts_sf <- tracts_sf %>%
  mutate(
    log_obesity = tracts$log_physact,  
    physact_base_lm_residuals = physact_base_lm$residuals,
    physact_base_sem_fitted = physact_base_sem$fitted.values,  
    physact_base_sem_residuals = physact_base_sem$residual
  )

ggplot(tracts_sf, aes(x = log_obesity, y = physact_base_sem_fitted)) + 
  geom_point()
```

@@

```{r logsums_function}

#' Calculate destination choice logsums from a distance matrix
#' 
#' @param d An $n\times p$ matrix with the distance from all tracts to
#'   all parks
#' @param sizes A p-length vector of park sizes
#' @param tweets A p-length vector of tweets at parks
#' @param betas A vector containing the size, distance, and tweet coefficients. 
#'   If we submit two variables the tweets are ignored.
#'   
#' @return An n-length vector containing the weighted log-sum based
#'   accessibility between a tract and all parks.
#' @details If we have n tracts and p parks, distances needs to be a 
#' 
calculate_park_logsums <- function(d, sizes, tweets = NULL,
                                   betas = c(-15,.00001, .001)){

@@

access_sdm_val <- optim(
  fn = notweets_logsum,  
  par = c(-0.42, 0.5), # distance, size 
  lower = c(-10, 0), # size coefficient must be positive 
  upper = c(0, 10), # distance coefficient must be negative 
  distances = log(distances), parks = parks, tracts = tracts, 
  model = physact_base_sem,
  method = "L-BFGS-B", 
  control = list(trace = 6))
```

```{r tweets_logsum, eval = FALSE}
tweets_logsum <- function(betas, distances, parks, tracts, model) {

  logsum <- calculate_park_logsums(
    d = distances,  sizes = parks$log_size, tweets = parks$log_tweets,
    betas = betas)

  tracts@data$tweets_ls <- (logsum - mean(logsum)) / sd(logsum)

  sdm <- update(model, . ~ . + tweets_ls)
  return(-logLik(sdm))
}

tweets_sdm_val <- optim(
  fn = tweets_logsum,  
  par = c(-0.42, 0.5, 1), # distance, size 
  lower = c(-10, 0, -10), # size coefficient must be positive 
  upper = c(0, 10, 10), # distance coefficient must be negative 
  distances = log(distances), parks = parks, tracts = tracts, 
  model = physact_base_sem,
  method = "L-BFGS-B", 
  control = list(trace = 6))

@@

```{r show_accessibility}
tracts@data <- tracts@data %>%
  mutate(
    access_ls = calculate_park_logsums(log(distances), parks$log_size,
                                       betas = c(-1.796, 1.0)),
    tweets_ls = calculate_park_logsums(log(distances), parks$log_size, parks$log_tweets,
                                       betas = c(-1.796, 1.0, 0.1)),
    access_ls = (access_ls - mean(access_ls) )/ sd(access_ls),
    tweets_ls = (tweets_ls - mean(tweets_ls) )/ sd(tweets_ls)
  )

tracts_sf <- tracts_sf %>%
  mutate(
    access_ls = tracts$access_ls,
    tweets_ls = tracts$tweets_ls
  )

pal <- colorQuantile("Spectral", c(tracts_sf$access_ls, tracts_sf$tweets_ls), 5)

leaflet(tracts_sf %>% st_transform(4326)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(group = "access", color = ~pal(access_ls)) %>%
  addPolygons(group = "access + tweets", color = ~pal(tweets_ls)) %>%
  addLayersControl(baseGroups = c("access", "access + tweets"))

@@

```{r modelchart}
physact_models <-  
screenreg()

```{r impacts_extractor}
#' Get the MC simulated impact coefficients and significance from a model
#' @param sdm An autoregressive lag (SAR or SDM) model object `lagsarlm`
#' @return A data frame with the effect (direct, indirect, total), variable, 
#'   simulated mean impact, and significance p-value
impacts_extractor <- function(sdm) {

  impacts_summary <- summary(
    impacts(sdm, tr=trMC,  R = 1000, useHESS = !is.logical(sdm$fdHess)), 
    zstats = TRUE)

  coef <- list(
    Direct = impacts_summary$direct_sum,
    Indirect = impacts_summary$indirect_sum,
    Total = impacts_summary$total_sum
  ) %>%
  lapply(function(s) {
    data_frame(
      var = names(s$statistics[,1]),
      impact = s$statistics[,1]
    )
  }) %>%
    bind_rows(.id = "effect")

  pval <- as_data_frame(impacts_summary$pzmat) %>%
    mutate(
     var = rownames(impacts_summary$pzmat)
    ) %>%
    gather("effect", "p-val", -var)

  left_join(coef, pval, by = c("effect", "var"))
}
```

```{r obesity_impacts}
obesity_impacts <- obesity_models %>%
  lapply(impacts_extractor) %>%
  bind_rows(.id = "model") %>%
  transmute(
    model, var, effect, 
    output = paste(round(impact, 5), 
                   gtools::stars.pval(`p-val`))
  ) %>%
  spread(model, output) 
```

# Physical Activity

We cannot necessarily assume that the logsum parameter will be constructed the
same way for physical activity as for obesity. (maybe?) But that's the assumption
we'll make for now.

```{r physact_estimation}
pa_base_sdm <- update(obese_base_sdm, log_physact ~ .)
pa_access_sdm <- update(access_sdm, log_physact ~ .)
pa_tweets_sdm <- update(tweets_sdm, log_physact ~ .)
```

```{r physact_models}
physact_models <- list(Base = pa_base_sdm, Access = pa_access_sdm, 
                       `Access + Tweets` = pa_tweets_sdm)
screenreg(physact_models, digits = 4)
```

```{r physact_impacts}
physact_impacts <- physact_models %>%
  lapply(impacts_extractor) %>%
  bind_rows(.id = "model") %>%
  transmute(
    model, var, effect, 
    output = paste(round(impact, 5), 
                   gtools::stars.pval(`p-val`))
  ) %>%
  spread(model, output) 
```

# Results
```{r impacts_table}
list("Obesity" = obesity_impacts, "Physical Activity" = physact_impacts) %>%
  bind_rows(.id = "Outcome") %>%
  filter(var %in% c("access_ls", "tweets_ls")) %>%
  dplyr::select(-var, -Base)  %>%
  gather("Measure", "Impact", -Outcome, -effect) %>%
  filter(!is.na(Impact))  %>%
  spread(effect, Impact)

>>>>

tracts@data <-  tbl_df(tracts@data) %>%
  dplyr::transmute(
    geoid = GEOID,
    county_fips = substr(GEOID, 3, 5),
    borough = case_when(
      county_fips == "081" ~ "Queens",
      county_fips == "047" ~ "Brooklyn",
      county_fips == "061" ~ "Manhattan",
      county_fips == "005" ~ "Bronx",
      county_fips == "085" ~ "Staten Island",
      TRUE ~ as.character(NA)
    ),
    obesity = OBESITY,
    physact = 100 - Phys_Act,
    log_obesity = log(OBESITY),
    log_physact = log(100 - Phys_Act),
    density = Pop_Density,
    fulltime = FulltimeWork, college = CollegeDeg,
    single = Single_Percent,
    black = PctBlack, asian = PctAsian, hispanic = PctHispanic,
    other = PctNative + PctPacific,
    Pct0to17, Pct18to29, Pct65plus,
    income = NYC_Income,
    lowincome = Income1 + Income2 + Income3,
    highincome = Income8 + Income9 + Income10
  )

tracts_sf <- tracts_sf %>% filter(GEOID %in% tracts@data$geoid)
```

@@

```{r W, echo=FALSE}
pop_weighted_centroids <- read_csv("data/CenPop2010_Mean_TR36.csv") %>%
  mutate(geoid = str_c(STATEFP, COUNTYFP, TRACTCE))
pop_weighted_centroids <- left_join(tracts@data, pop_weighted_centroids, by = "geoid") %>%
  dplyr::select(geoid, LONGITUDE, LATITUDE) %>%
  st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
  st_transform(2263)

tracts.dnn <- dnearneigh(gCentroid(as(pop_weighted_centroids, "Spatial"), byid = TRUE),
                         0, 1.8 * 5280)
dists <- nbdists(tracts.dnn, gCentroid(tracts, byid = TRUE))
dists.inv <- lapply(dists, function(x) 1 / x)
W <- nb2listw(neighbours = tracts.dnn, glist = dists.inv,
              zero.policy = TRUE, style = "W")
trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
```

@@

```{r base_formula, message=FALSE, include=FALSE}
base_formula <- formula(
  ~ log(density) + highincome + lowincome +
    fulltime + college + single +
    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity
    black + asian + hispanic + other)
```

@@

```{r spatial_selection}
physact_base_lm <- lm(update(base_formula, physact ~ .), data = tracts)
physact_base_sar <- lagsarlm(update(base_formula, physact ~ .),
                           data = tracts@data, listw = W, zero.policy = TRUE)
physact_base_sem <- errorsarlm(update(base_formula, physact ~ .),
                             data = tracts@data, listw = W, zero.policy = TRUE)
physact_base_sdm <- lagsarlm(update(base_formula, physact ~ .),
                           data = tracts@data, listw = W, zero.policy = TRUE,
                           type = "mixed")
test_sdmsem <- lmtest::lrtest(physact_base_sem, physact_base_sdm)
```

@@

```{r physact_models, results='asis'}
physact_models <- list("OLS" = physact_base_lm, "SAR" = physact_base_sar, 
                       "SEM" = physact_base_sem, "SDM" = physact_base_sdm)
htmlreg(physact_models, digits = 5)
```

@@

```{r residuals}
tracts_sf <- tracts_sf %>%
  mutate(
    physact = tracts$physact,  
    physact_base_lm_residuals = physact_base_lm$residuals,
    physact_base_sem_fitted = physact_base_sem$fitted.values,  
    physact_base_sem_residuals = physact_base_sem$residual
  )

ggplot(tracts_sf, aes(x = physact, y = physact_base_sem_fitted)) + 
  geom_point()
```

@@

```{r logsums_function}
#' Calculate destination choice logsums from a distance matrix
#'
#' @param d An $n\times p$ matrix with the distance from all tracts to
#'   all parks
#' @param sizes A p-length vector of park sizes
#' @param tweets A p-length vector of tweets at parks
#' @param betas A vector containing the size, distance, and tweet coefficients.
#'   If we submit two variables the tweets are ignored.
#'
#' @return An n-length vector containing the weighted log-sum based
#'   accessibility between a tract and all parks.
#' @details If we have n tracts and p parks, distances needs to be a
#'
calculate_park_logsums <- function(d, sizes, tweets = NULL,
                                   betas = c(-0.4,0.2,0.001)){

@@

access_sdm_val <- optim(
  fn = notweets_logsum,  
  par = c(-1.79, 0.5), # distance, size 
  lower = c(-10, 0), # size coefficient must be positive 
  upper = c(0, 10), # distance coefficient must be negative 
  distances = log(distances), parks = parks, tracts = tracts, 
  model = physact_base_sem,
  method = "L-BFGS-B", 
  control = list(trace = 6))
```

The final estimated values are $\beta_d$= `r access_sdm_val$par[1]`,
$\beta_s$= `r access_sdm_val$par[2]`, though this presents problems from 
two perspectives. First, we assert that the size of a park is not 
immaterial in its attractiveness. Second, ignoring the size of a park results in
tracts immediately adjacent to Central Park with poor accessibility, which is
unreasonable. So we assert a value of 0.8.

```{r show_accessibility}
tracts@data <- tracts@data %>%
  mutate(
    access_ls = calculate_park_logsums(log(distances), parks$log_size,
                                       betas = c(-1.96805, 0.8)),
    estimd_ls = calculate_park_logsums(log(distances), parks$log_size,
                                       betas = c(-1.96805, 0.0)),
    tweets_ls = calculate_park_logsums(log(distances), parks$log_size,
                                       parks$log_tweets,
                                       betas = c(-1.96805, 0.8, 0.1)),
    access_ls = (access_ls - mean(access_ls) )/ sd(access_ls),
    estimd_ls = (estimd_ls - mean(estimd_ls) )/ sd(estimd_ls),
    tweets_ls = (tweets_ls - mean(tweets_ls) )/ sd(tweets_ls)
  )

tracts_sf <- tracts_sf %>%
  select(-access_ls, -estimd_ls, -tweets_ls) %>%
  left_join(tracts@data %>% select(geoid, access_ls, estimd_ls, tweets_ls), 
            by = c("GEOID" = "geoid"))

pal <- colorQuantile("Spectral", c(tracts_sf$access_ls, tracts_sf$estimd_ls, 
                                   tracts_sf$tweets_ls), 5)

leaflet(tracts_sf %>% st_transform(4326)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(group = "access", color = ~pal(access_ls)) %>%
  addPolygons(group = "maximum likelihood", color = ~pal(estimd_ls)) %>%
  addPolygons(group = "access + tweets", color = ~pal(tweets_ls)) %>%
  addLayersControl(baseGroups = c("access", "maximum likelihood", 
                                  "access + tweets"))

@@

Now that we have the dependent variable, we can look at the relationship between
access and physical activity.

@@
